---
title: Dynamic Energy Budget (DEB) theory summary notes
author: |
 | Matthew Malishev^1^*  
 |
 | _^1^ Department of Biology, Emory University, 1510 Clifton Road NE, Atlanta, GA, USA, 30322_
#bibliography: ~/library.bib
fontsize: 10
geometry: margin=1in
documentclass: article
linkcolor: pink
urlcolor: blue
citecolor: red
output:
  pdf_document:
    includes:
      in_header: # add .tex file with header content
    highlight: tango
    template: null
    toc: yes
    toc_depth: 4
    number_sections: false
    fig_width: 5
    fig_height: 6
    fig_caption: true
    df_print: tibble 
    citation_package: biblatex # natbib
    latex_engine: xelatex #pdflatex # lualatex
    keep_tex: true # keep .tex file in dir 
  html_document:
    highlight: tango
    code_folding: hide
    toc: yes
    toc_depth: 4
    number_sections: no
    toc_float: yes
  word_document:
    highlight: tango
    keep_md: yes
    pandoc_args: --smart
    #reference: mystyles.docx
    toc: yes
inludes:
  before_body: #before_body.tex
subtitle: 
tags:
- nothing
- nothingness
params: 
  dir: "/Users/malishev/Documents/Melbourne Uni/Programs/DEB/deb"
  date: !r Sys.Date()
  version: !r getRversion()
  email: "matthew.malishev@gmail.com"
  doi: https://github.com/darwinanddavis/DEB_notes
classoption: portrait
# ^['https://github.com/darwinanddavis/DEB_notes'] # footnote
vignette: >
  %\VignetteIndexEntry{DEB notes}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
# library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = FALSE, cache = FALSE}
options(width=100)
knitr::opts_chunk$set(
 eval = TRUE, # run all code
 # echo = FALSE, # show code chunks in output 
 comment = "",
 tidy.opts=list(width.cutoff=100), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

setwd(paste0(params$dir,"/")) # for running just in R not knitr
```

\  

Date: `r params$date`  
R version: `r params$version`  
*Corresponding author: `r params$email`  

\newpage  

## List of parameters and variables  

|Parameter|Definition|  
|:------:|:------|  
| $E$ | energy reserve |  
| $[E]$ | energy reserve per volume (reserve density) |  
| $e$ | scaled energy reserve |
| $[E_M]$ | maximum reserve density |    
| $f$ | functional feeding response |  
| $L$ | structural length |   
| $\dot{p}_{AM}$ | maximum assimilation rate |  
| $\dot{v}$ | energy conductance |     


\newpage  

## Overview   

Summarised notes from DEB workshops, telecourses, lectures, and discussions.  

### Reserve mobilisation  
Conductance determines mobilisation rate from reserve to structure  
The larger the surface area of reserve, the more mobilisation is possible and thus faster maintenance and growth due to more surface area.  
- surface area scales slower than volume-specific energy flows.     

Reserve dynamics
f = 1 (max feeding rate)  
$$
\frac
  {d E}
  {d t}  
  = \frac
  {f\{\dot{p}_{AM}\}}
  {L}  
  - \frac  
  {\dot{v}[E]}  
  {L}   
$$

$[E_{M}]$ = max reserve. Reserve doesn’t change.

$$
= \frac
  {\{\dot{p}_{AM}\}}
  {L} - 
  \frac
  {\dot{v}[E_{M}]}
  {L}    
$$
$$
\therefore 
[E_{M}]
= \frac
  {\{\dot{p}_{AM}\}}
  {\dot{v}}
$$
Scaled reserve
$$
e =
\frac
{[E]}
{[E_{M}]}  
$$

$$
\frac
{de}
{dt}
= \frac
{[E] /[E_{M}]}
{dt} 
= \frac
{f\dot{v}}
{L}  - 
\frac
{e\dot{v}}
{L}     
$$  
$$
= \frac
{\dot{v} (f – e)}
{L} 
$$

Under steady state, reserve doesn’t change   
$$
0 =
\frac
{\dot{v} (f – e)}
{L}
\ \ \ \
\text{or}
\ \ \ \ f = e
$$

### Length  
  
Getting maximum length $L_{m}$      

$$
\frac
{dV}
{dt}
= V\dot{r}  
$$

Can rewrite $r$ using scaled reserve $e$

$$
\dot{r} = 
\dot{v}
\frac
{\frac
{e}
{L} - 
(1 + \frac
{L_{T}} 
{L}
) /
{L_{m}}}
{e + g}
$$
Getting $L_{m}$    

$$
\frac
{dV}
{dt} = 
V\dot{r}
$$

To find $V_{m} = Lm^{3}$, set $f = 1$ and $\frac{dV}{dt} = 0$, then solve for $V = V_{m}$    
  
$$
L_{m}
=
\frac
{\kappa\{\dot{p}_{Am}\}}
{[\dot{p}_{M}]}
$$
  
### Weak homeostasis  
Structural isomorphy implies weak homeostasis  
Weak homeostasis depends on ratio of reserve to structure $\frac{d[E]}{dt}$  





